---
title: "Constructing the Eikosany"
output: rmarkdown::html_vignette
bibliography: eikosany.bibtex
vignette: >
  %\VignetteIndexEntry{Constructing the Eikosany}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(eikosany)
```

## Introduction
The primary reference for this package is [@narushima2019microtonality]. Some
definitions for context:

- _Microtonality_: Technically, _microtonality_ refers to musical scales with
more notes than the standard 12 notes of the Western equal-tempered scale.
However, the definition has been broadened somewhat to include scales with 12
or fewer notes, as long as they aren't identical to the Western  standard ones.
- _Harmonic spectrum_: A note has a _harmonic spectrum_ if the spectrum
consists of a fundamental frequency plus overtones that are integer multiples
of the fundamental.
- _Just intonation_: _Just intonation_ scales are designed to have intervals
that are consonant when played on musical instruments whose spectrum is
harmonic. The Eikosany is a just intonation scale. See [@sethares2013tuning]
for the mathematics of harmonic spectra and just intonation.

```{r}
library(gmp)
library(data.table)
big1 <- as.bigq(1)
big2 <- as.bigq(2)
big15 <- as.bigq(15)

octave <- function(x) {
  w <- as.bigq(x)
  
  ix <- (w > big2)
  while (any(ix)) {
    w[ix] <- w[ix] / big2
    ix <- (w > big2)
  }
  
  ix <- (w < big1)
  while (any(ix)) {
    w[ix] <- w[ix] * big2
    ix <- (w < big1)
  }

  return(w)
}
```

```{r}
harmonics <- c(1, 3, 5, 7, 9, 11)
eikosany <- as.data.table(t(combn(harmonics, 3)))
eikosany <- eikosany[, .(
  cps_label = paste(V1, V2, V3, sep = "*"), product = V1 * V2 * V3
  )]
eikosany <- eikosany[, .(
  cps_label, ratio = octave(as.bigq(product) / big15)
  )]
eikosany <- eikosany[, .(
  cps_label, ratio = as.character(ratio), num_ratio = as.numeric(ratio)
  )]
setkey(eikosany, num_ratio)
eikosany <- eikosany[, .(
  cps_label, ratio, num_ratio, midi_note_number = 60:79
  )]
print(eikosany)
```

```{r}
tetrads <- t(combn(harmonics, 4))
for (it in 1:nrow(tetrads)) {
  tx <- tetrads[it, ]
  print(paste(tx, collapse = "|"))
  others <- setdiff(harmonics, tx)
  notes <- c()
  inv_notes <- c()

  for (nx in 1:length(tx)) {
    note <- paste(sort(union(others, tx[nx])), collapse = "*")
    notes <- union(notes, note)
    inv_note <- paste(sort(setdiff(tx, tx[nx])), collapse = "*")
    inv_notes <- union(inv_notes, inv_note)
  }
  
  print(notes)
  print(inv_notes)
  
}
```

## References
