---
title: "Constructing the Eikosany"
output: rmarkdown::html_vignette
bibliography: eikosany.bibtex
vignette: >
  %\VignetteIndexEntry{Constructing the Eikosany}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(eikosany)
```

## Introduction
The primary reference for this package is [@narushima2019microtonality]. Some
definitions for context:

- _Microtonality_: Technically, _microtonality_ refers to musical scales with
more notes than the standard 12 notes of the Western equal-tempered scale.
However, the definition has been broadened somewhat to include scales with 12
or fewer notes, as long as they aren't identical to the Western  standard ones.
- _Harmonic spectrum_: A note has a _harmonic spectrum_ if the spectrum
consists of a fundamental frequency plus overtones that are integer multiples
of the fundamental.
- _Just intonation_: _Just intonation_ scales are designed to have intervals
that are consonant when played on musical instruments whose spectrum is
harmonic. The Eikosany is a just intonation scale. See [@sethares2013tuning]
for the mathematics of harmonic spectra and just intonation.

```{r}
library(sets)

harmonics <- set("A", "B", "C", "D", "E", "F")
subharmonics <- set("1/A", "1/B", "1/C", "1/D", "1/E", "1/F")
eikosany <- set_combn(harmonics, 3)
tetrads <- set_combn(harmonics, 4)
subtetrads <- set_combn(subharmonics, 4)

for (t in tetrads) {
  others <- harmonics - t
  print(others)
  notes <- set()

  for (tx in t) {
    note <- set_union(others, tx)
    notes <- set_union(notes, set(note))
  }
  
  print(notes)
  
}








```
## References
