---
title: "Constructing the Eikosany"
output: rmarkdown::html_vignette
bibliography: eikosany.bibtex
vignette: >
  %\VignetteIndexEntry{Constructing the Eikosany}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(eikosany)
```

## Motivation

On February 4, 2001, composer Iannis Xenakis passed away. I've been a fan of
experimental music, especially _musique concrète_, algorithmically composed
music, microtonal music, and other _avant-garde_ genres since I was an
undergraduate. Xenakis was one of the major figures in algorithmic
composition.

Reading the first edition of _Tuning, Timbre, Spectrum, Scale_
[@sethares1998tuning] rekindled my appreciation for the microtonal music of
Harry Partch. And so, armed with copies of @sethares1998tuning,
_Formalized Music_ [@xenakis1992formalized], and _Genesis of a Music_
[@partch1979genesis], I embarked on a path that led to
_When Harry Met Iannis_ [@Borasky2021a].

_When Harry Met Iannis_ was premiered at a microtonal music festival in El
Paso, Texas in late October, 2001. The
[Bandcamp version](https://algocompsynth.bandcamp.com/album/when-harry-met-iannis)
is essentially identical to that version; the source code is on GitHub at
<https://github.com/AlgoCompSynth/when-harry-met-iannis>.

At the festival, I met a number of composers who were working in microtonal
and just intonation, and one name kept coming up: Erv Wilson. Wilson was a
theoretician who developed keyboards, scales and tuning systems that
several composers were using at the time, and are still using today. Terumi
Narushima's _Microtonality and the Tuning Systems of Erv Wilson_
[@narushima2019microtonality] is a comprehensive documentation of Wilson's
work and is the basis for the code in this package.

The 20th anniversary of Xenakis' passing and of _When Harry Met Iannis_
occured in my second year of virtual isolation because of COVID-19. During
2021, I acquired two synthesizers that are capable of mapping the keyboards
to arbitrary microtonal scales: an
[Ashun Sound Machines Hydrasynth Desktop](https://www.ashunsoundmachines.com/hydrasynth-desk), and a
[Korg Minilogue XD](https://www.korg.com/us/products/synthesizers/minilogue_xd/).

The Hydrasynth ships with the tuning tables for many of Erv Wilson's scales
already in the firmware. For the Minilogue XD, the user can load up to six
custom scales with a software librarian program. In either case, though, I'm
not a keyboard player, and even if I was, the chords in the Wilson scales
are not where they would be on a conventional keyboard; the synthesizer has
remapped them. So I needed a tool to generate MIDI files to play the chords.
And that's what this package does.

## Computing Combination Product Set scales

```{r}

# the defaults yield the 1-3-5-7-9-11 eikosany
eikosany <- scale_table()
print(eikosany[["scale_table"]])
print(eikosany[["interval_matrix"]])

# the 1-7-9-11-13 dekany
dekany_harmonics <- c(1, 7, 9, 11, 13)
dekany_choose <- 2
dekany_1_3_7_9_11_13 <-
  scale_table(dekany_harmonics, dekany_choose)
print(dekany_1_3_7_9_11_13[["scale_table"]])
print(dekany_1_3_7_9_11_13[["interval_matrix"]])

# We might want to print out sheet music for a conventional keyboard
# player, since the synthesizer is mapping MIDI note numbers to pitches.
# We assume at least a 37-key synthesizer with middle C on the left.
# so the largest CPS scale we can play is a 35-note "35-any", made from
# seven harmonics taken three at a time.
harmonics_35 <- c(1, 3, 5, 7, 9, 11, 13)
choose_35 <- 3
any_35 <-
  scale_table(harmonics_35, choose_35)
print(any_35[["scale_table"]])
print(any_35[["interval_matrix"]])

```

## Computing chords - the tetrads of the 1-3-5-7-9-22 eikosany

```{r}
# the defaults yield the 1-3-5-7-9-11 eikosany
eikosany_scale <- scale_table()[["scale_table"]]
print(eikosany_scale)

# the defaults yield the tetrads of the 1-3-5-7-9-11 eikosany
eikosany_chords <- chord_table(eikosany_scale)
print(eikosany_chords)

```

## References
