---
title: "Combination Product Sets on the Korg Minilogue XD"
output: rmarkdown::html_vignette
bibliography: eikosany.bibtex
vignette: >
  %\VignetteIndexEntry{Combination Product Sets on the Korg Minilogue XD}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(eikosany)
```

## The Korg Minilogue XD
The [Korg Minilogue
XD](https://www.korg.com/us/products/synthesizers/minilogue_xd/)
is a digital-controlled four-voice analogue synthesizer with an added digital
multi-engine that provides digital signal processing capabilities -
oscillators, filters, and effects. The Minilogue XD comes in two form
factors - a synthesizer with a 37-key keyboard and a module without the
keyboard.

There is a [Sound
Librarian](https://www.korg.com/us/products/synthesizers/minilogue_xd/librarian_contents.php)
for managing the sound programs and multi-engine software in the Minilogue
XD, and a [Software Development Kit - `logue
SDK`](https://www.korg.com/us/products/synthesizers/minilogue_xd/sdk.php)
for developing multi-engine software.

The Minilogue XD and Minilogue XD Module can be microtuned to arbitrary
scales in two ways:

1. user octaves, and
2. user scales [@Korg2019a].

Both methods can be implemented directly on the device or uploaded to the
device with the Sound Librarian.

The *user octave* method can be used for any scale with 12 or fewer notes per
octave. Each octave in the whole range of the instrument is tuned to the same
settings, expressed as offsets in cents from the standard 12 equal divisions
of the octave (12EDO) tuning. Thus, you can tune a Minilogue XD or Minilogue
XD Module to a Hexany [@narushima2019microtonality, 151] or Dekany
[@narushima2019microtonality, 160] by user octaves, using the output of
`create_scale_table`.

### 1-3-5-7 Hexany
```{r}
hexany <- create_scale_table(harmonics = c(1, 3, 5, 7), choose = 2)
print(hexany)

```

### 1-3-5-7-9 2)5 Dekany
```{r}
dekany <- create_scale_table(harmonics = c(1, 3, 5, 7, 9), choose = 2)
print(dekany)

```


## User scales and Eikosanies
*User scales* retune the entire range of the device - MIDI note numbers zero
through 127 - to any pitches, represented as an offset in cents from any
standard 12EDO key. Since an Eikosany [@narushima2019microtonality, 161]
has 20 notes to the octave, you can't retune a Minilogue XD / XD Module to
an Eikosany with user octaves - you have to use user scales.

## References
