---
title: "Introduction to Combination Product Sets"
output: rmarkdown::html_vignette
bibliography: eikosany.bibtex
vignette: >
  %\VignetteIndexEntry{Introduction to Combination Product Sets}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(eikosany)
```

## Alternative Tunings Overview

Unless you're into alternative tunings, your first question is probably
"Alternative to what?" Of course, that depends on what culture you call
home. But for most Americans and other native English speakers, most
Europeans, and mass-market popular music performers and audiences nearly
everywhere, music is tuned to 12-tone equal temperament, abbreviated as
"*12-TET*". You will also see this tuning referred to as "*12-EDO*",
which stands for "twelve equal divisions of the octave."

So what are the alternatives? First, there are tunings from music that
evolved in cultures other than the ancient Greek / Italian / French /
German mainstream (with nationalist enhancements) that now use 12-EDO.
Notable examples are Indonesian
[*gamelan*](https://en.wikipedia.org/wiki/Gamelan "Gamelan on Wikipedia"),
South Indian [*carnatic*
music](https://en.wikipedia.org/wiki/Carnatic_music "Carnatic music on Wikipedia"),
and various
[*makams*](https://en.wikipedia.org/wiki/List_of_makams "List of makams on Wikipedia")
from [Ottoman
music](https://en.wikipedia.org/wiki/Ottoman_music "Ottoman music on Wikipedia").

The other alternative tunings have been derived by theorists, composers,
and performers over time, first on the path from ancient Greek music to
12-EDO claiming the dominant market share it has today, and since then,
as a counter to said market share. There are two main types of such
alternative tunings in use today.

The first is equal divisions of the octave with more than twelve notes,
generally called "N-TET" or "N-EDO". For example, quarter-tone scales
are *24-EDO*. Musicians create scales of this kind for a variety of
reasons, most often to provide more harmonic possibilities without
giving up the ability to modulate to different keys. Two such scales
that have achieved serious usage are
[*19-EDO*](https://en.wikipedia.org/wiki/19_equal_temperament "19 equal temperament on Wikipedia")and
[*31-EDO*](https://en.wikipedia.org/wiki/31_equal_temperament "31 equal temperament on Wikipedia").
