---
title: "Introduction to Combination Product Sets"
output: rmarkdown::html_vignette
bibliography: eikosany.bibtex
vignette: >
  %\VignetteIndexEntry{Introduction to Combination Product Sets}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(eikosany)
```

## Alternative Tunings Overview

Unless you're into alternative tunings, your first question is probably
"Alternative to what?" Of course, that depends on what culture you call
home. But for most Americans and other native English speakers, most
Europeans, and mass-market popular music creators and audiences nearly
everywhere, music is tuned to 12-tone equal temperament, abbreviated as
"*12-TET*". You will also see this tuning referred to as "*12-EDO*",
which stands for "twelve equal divisions of the octave."

### Other cultures

So what are the alternatives? First, there are tunings from music that
evolved in cultures other than the ancient Greek / Italian / French /
German mainstream (with nationalist enhancements) that now uses 12-EDO.
Notable examples are Indonesian
[*gamelan*](https://en.wikipedia.org/wiki/Gamelan "Gamelan on Wikipedia"),
South Indian [*carnatic*
music](https://en.wikipedia.org/wiki/Carnatic_music "Carnatic music on Wikipedia"),
and various
[*makams*](https://en.wikipedia.org/wiki/List_of_makams "List of makams on Wikipedia")
from [Ottoman
music](https://en.wikipedia.org/wiki/Ottoman_music "Ottoman music on Wikipedia").

You can find detailed analyses of gamelan tunings, as well as Thai
classical tunings and numerous theoretical and practical tunings in
[@sethares2013tuning]. The first edition of this book
[@sethares1998tuning] is the book that sent me down the alternate
tunings rabbit hole in 2001.

### Theorists, composers and performers

The other alternative tunings have been derived by theorists, composers,
and performers over time, first as steps along the same path from
ancient Greek music to 12-EDO claiming the dominant market share it has
today, and since then, as a counter to said market share. There are two
main types of such alternative tunings in use today.

The first is equal divisions of the octave with more than twelve notes,
generally called "N-TET" or "N-EDO". For example, quarter-tone tunings
are *24-EDO*. Musicians create tunings of this kind for a variety of
reasons, most often to provide more harmonic possibilities without
giving up the ability to modulate to different keys.

Two such tunings that have achieved serious usage are
[*19-EDO*](https://en.wikipedia.org/wiki/19_equal_temperament "19 equal temperament on Wikipedia")
and
[*31-EDO*](https://en.wikipedia.org/wiki/31_equal_temperament "31 equal temperament on Wikipedia").
If you're curious about such tunings, function `et_scale_table` in this
package can create scale tables for them, and function `keyboard_map`
can create a mapping for such a scale to a synthesizer keyboard.

Here's what the 19-EDO scale table and keyboard map look like:

```{r}
data("edo19_names")
edo19_scale_table <- et_scale_table(edo19_names)
print(edo19_scale_table)
edo19_keyboard_map <- keyboard_map(edo19_scale_table)
print(edo19_keyboard_map)
```

### Just intonation

The other main class of derived tunings is called [*just
intonation*](https://en.wikipedia.org/wiki/Just_intonation "Just intonation on Wikipedia").
Just intonation attempts to create perfect harmonies by dividing the
octave into unequal intervals using ratios of small integers. For
example, the just perfect fifth is 3/2, the just perfect fourth is 4/3,
and the just major third is 5/4.

If you've heard a barbershop quartet, you've heard just intonation.
Combination product sets are a form of just intonation.

## Combination Product Sets

Combination product sets [@narushima2019microtonality, chapter 6] are
just one of theoretician Erv Wilson's tuning constructs. The current
release, v0.5.0, is focused on them, and the ability to play music with
them on synthesizers that can be retuned, such as the [Korg Minilogue
XD](https://www.korg.com/us/products/synthesizers/minilogue_xd/ "Korg Minilogue XD web page"),
the [Aodyo Anyma
Phi](https://aodyo.com/anyma-phi/ "Aodyo Anyma Phi web page"), and the
[Ashun Sound
Machines](https://www.ashunsoundmachines.com/ "Ashun Sound Machines home page")
Hydrasynth.

## References
