---
title: "Introduction to Combination Product Sets"
output: rmarkdown::html_vignette
bibliography: REFERENCES.bib
vignette: >
  %\VignetteIndexEntry{Introduction to Combination Product Sets}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(eikosany)
```

## Alternative Tunings Overview

Unless you're into alternative tunings, your first question is probably
"Alternative to what?" Of course, that depends on what culture you call
home. But for most Americans and other native English speakers, most
Europeans, and mass-market popular music creators and audiences nearly
everywhere, music is tuned to 12-tone equal temperament, abbreviated as
"*12-TET*". You will also see this tuning referred to as "*12-EDO*",
which stands for "twelve equal divisions of the octave."

So what are the alternatives? There are two main sources of alternative
tunings:

-   Other cultures, and

-   Music theorists / composers / instrument builders / performers
    within the ancient Greek / Italian / French / German mainstream
    (with nationalist enhancements!) that now uses 12-EDO. For brevity
    we'll call those "Western theoretical tunings".

### Other cultures

Notable examples of tunings from other cultures are Indonesian
[*gamelan*](https://en.wikipedia.org/wiki/Gamelan "Gamelan on Wikipedia")
tuning, South Indian [*carnatic*
music](https://en.wikipedia.org/wiki/Carnatic_music "Carnatic music on Wikipedia"),
and various
[*makams*](https://en.wikipedia.org/wiki/List_of_makams "List of makams on Wikipedia")
from [Ottoman
music](https://en.wikipedia.org/wiki/Ottoman_music "Ottoman music on Wikipedia").
You can find detailed analyses of gamelan tunings, as well as Thai
classical tunings and numerous theoretical and practical tunings in
[@sethares2013tuning]. The first edition of this book
[@sethares1998tuning] is the book that sent me down the alternate
tunings rabbit hole in 2001.

### Western theoretical tunings

Numerous tunings and temperaments have derived by theorists, composers,
and performers over time. This happened first as steps along the path
from ancient Greek music to 12-EDO claiming the dominant market share it
has today, and later as a counter to said market share.

In some cases, instrument builders constructed moderately complex
instruments that used these tunings. It was the desire to mass-produce
musical instruments, especially the piano, that led to the widespread
adoption of 12-EDO as the standard.

There are two main types of such alternative tunings in use today:

-   Equal divisions of a period, usually the octave, and

-   Just intonation.

### Brief guide to tuning and scales

Before moving onward, we need to define some concepts. *Pitches* are
(sort of) the musical name for frequencies. *Frequencies* are the rate
at which some string or air column or reed or membrane or loudspeaker or
other physical object vibrates.

Frequencies are designated in cycles per second, called *Hertz* and
abbreviated *Hz*. The typical range of human hearing is 20 - 20,000
Hertz, but not all of that range is commonly used in music.

A *tuning* describes a set of pitches a composer selects from in writing
a piece, often dictated by the instruments that will be used to play it.
Tunings usually consist of a set of repetitions of a *scale.* A scale is
a set of pitches in ascending order of frequency.

The most common tuning of this type is equal divisions of the octave
with more than twelve notes, generally called "N-TET" or "N-EDO". For
example, quarter-tone tunings are *24-EDO*. Musicians create tunings of
this kind for a variety of reasons, most often to provide more harmonic
possibilities without giving up the ability to modulate to different
keys.

Two such tunings that have achieved serious usage are
[*19-EDO*](https://en.wikipedia.org/wiki/19_equal_temperament "19 equal temperament on Wikipedia")
and
[*31-EDO*](https://en.wikipedia.org/wiki/31_equal_temperament "31 equal temperament on Wikipedia").
If you're curious about such tunings, function `et_scale_table` in this
package can create scale tables for them, and function `keyboard_map`
can create a mapping for such a scale to a synthesizer keyboard.

Here's what the 19-EDO scale table and keyboard map look like:

```{r}
data("edo19_names")
edo19_scale_table <- et_scale_table(edo19_names)
print(edo19_scale_table)
edo19_keyboard_map <- keyboard_map(edo19_scale_table)
print(edo19_keyboard_map)
```

### Just intonation

The other main class of derived tunings is called [*just
intonation*](https://en.wikipedia.org/wiki/Just_intonation "Just intonation on Wikipedia").
Just intonation attempts to create perfect harmonies by dividing the
octave into unequal intervals using ratios of small integers. For
example, the just perfect fifth is 3/2, the just perfect fourth is 4/3,
and the just major third is 5/4.

If you've heard a barbershop quartet, you've heard just intonation.
Combination product sets are a form of just intonation.

## Combination Product Sets

Combination product sets [@narushima2019microtonality, chapter 6] are
just one of theoretician Erv Wilson's tuning constructs. The current
release, v0.5.0, is focused on them, and the ability to play music with
them on synthesizers that can be retuned, such as the [Korg Minilogue
XD](https://www.korg.com/us/products/synthesizers/minilogue_xd/ "Korg Minilogue XD web page"),
the [Aodyo Anyma
Phi](https://aodyo.com/anyma-phi/ "Aodyo Anyma Phi web page"), and the
[Ashun Sound
Machines](https://www.ashunsoundmachines.com/ "Ashun Sound Machines home page")
Hydrasynth.

## References
